### 2019-08-29 concepts.tex
fix through three maybe

Delta-debugging \cite{DD}\footnote{We use delta-debugging to stand for
  all test reduction algorithms, even those \cite{CReduce,tstl} that do not use the exact
  binary-search that distinguishes delta-debugging \emph{per se}; the
  differences are immaterial for our purposes.}  is a widely used method for reducing the
size of failing tests, making them easier to understand and debug.
The core idea of delta-debugging is to take a test that satisfies some
property (usually the property is ``it fails in a particular way'')
and produce a smaller test that satisfies the same property.
Delta-debugging, as proposed by Zeller and Hildebrandt, relies on a
modified binary search to produce this smaller test, though the
following discussion relies only on the general structure of
delta-debugging, and also applies to the greedy algorithms used to
reduce failing tests in some work \cite{CReduce,onetest}.  At a high
level delta-debugging and other test reduction/shrinking algorithms
can be seen as following a general algorithmic structure.  Given a
test case, $T_{\mathit{INIT}} : \mathit{test}$ and a predicate $\mathit{PRED}: \mathit{test} \rightarrow \mathit{bool}$

@@

Delta-debugging in the context of detecting nondeterminism has two
purposes.  One is simply the usual goal of reducing the size of a
test.  Identifying the cause of nondeterminism may be very easy in a
test consisting of ten library function calls (it is one of these ten
calls), but very difficult in a test consisting of a hundred library
function calls.  This is no different than the common use of
delta-debugging.  However, in horizontal nondeterminism detection,
delta-debugging also tends to \emph{change the probability of
  nondeterministic behavior}; this can be both harmful and
beneficial.  The behavior is part of a more general (and, to our
knowledge, not previously investigated) issue:  using delta-debugging
to minimize a test with respect to a predicate that only holds with a
certain probability (the predicate itself is nondeterministic).  Note
that while most descriptions of delta-debugging discuss reducing a
test with respect to \emph{test failure} (hence the term
``delta-\emph{debugging}'') we adopt the wider view of delta-debugging
as reducing the length of a test while maintaining that an arbitrary
predicate remains true of the shorter versions of the test.  The
traditional predicate is ``the test fails'' but other predicates,
especially those related to code-coverage, can also be useful
\cite{icst2014,stvrcausereduce,NonAdeq}.  Our contribution here is to consider
what happens when the predicate to be evaluated is not deterministic.
``The test behaves flakily'' is a simple, important, example of such a predicate.

In \emph{monotonic} probabilistic delta-debugging, removing a part of a test cannot
increase the probability of the predicate holding.  This is fairly
common.  In these cases, when we use a reduction algorithm to reduce $t$
to $r$, $P(pred(r)) \leq P(pred(t))$.  In \emph{non-monotonic} cases,
however, there is no such upper bound.  Removing a step in $t$ may
increase the probability that $t$ behaves nondeterministically.

The probabilistic predicate of interest for determinism is basically of
the form:  {\bf the test will exhibit at least two different behaviors in $S$
runs, with probability $P$}.  This predicate may be monotonic or
non-monotonic, depending on the causes of the nondeterminism detected.

The only absolutely necessary changes required to use standard
delta-debugging implementations in nondeterminism detection
are simply removal of some ``sanity checks'' in the code:  many
implementations (including the Python code provided by Zeller) assert
that the predicate holds on the original test.  In probabilistic
settings, this may not be true, and we may even be trying to find a
subset where a predicate that is very far from holding on the original test
holds (in a non-monotonic case where we aim to increase the probability
of nondeterminism).

@@

However, simply using delta-debugging off the shelf with a predicate
like $P(fail) > 0.3 \wedge P(fail) < 1.0$, to force a test to be flaky, and force it to fail
sufficiently often to be used in debugging, will often produce
surprising and unfortunate results.  In many cases,
delta-debugging will indeed reduce the large test to a small subset.
And, in a technical sense, delta-debugging will work:  it will never
convert a nondeterministic test to a completely deterministic test,
because the reduced test that delta-debugging returns is always one
where the predicate of interest has been seen to evaluate to true.
However, if you run the resulting test, it will, in many cases, have a
$P(fail)$ that is much, much smaller than
0.3, perhaps as low as 0.01.  Why?

The problem is analogous to the problem of publication bias in
scientific fields \cite{ahmed2012assessment}.  A predicate like
$P(fail) > 0.3 \wedge P(fail) < 1.0$ cannot be evaluated by
determining the true probability of failure; rather, the test must be
run some concrete number of times, and the number of failures counted.
However, even if the number of samples $N$ is large, there is some
probability (based on the sample size) of a result that diverges
significantly from the actual probability of failure.  If the
predicate were run only once, and the number of samples reasonably
large, this would not matter.  However, by their nature,
delta-debugging and other test reduction algorithms explore a search
space that often contains hundreds or even thousands or millions of
candidate tests.  The predicate is evaluated on each of these, and so
even with large $N$, it is extremely likely that some evaluation will
produce a very poor estimate of $P(fail)$.  If such an evaluation
causes the predicate to appear to hold for a test, delta-debugging is
``stuck'' with the error, because (to our knowledge) no test reduction
algorithms allow backtracking.  After such a mistake, finding further
reductions will become harder, but may still be possible due to the
same source of errors:  the number of experiments run is far larger
than the probability of an incorrect result.

@@

In scientific literature, the most frequently proposed solution is the
use of replications:  repeated runs of ``successful'' experiments to
minimize the probability that a result is a fluke due to publication
bias.  One way to produce this effect would be to allow
delta-debugging to backtrack if the probabilities observed in predicate
evaluations suddenly exhibit a strong discontinuity, a kind of ``paper
retraction'' based on near-replications.  However, this requires
modifying the delta-debugging implementations, which is difficult and
sometimes not really feasible (e.g., few users of afl-fuzz \cite{aflfuzz} will wish
to change its C code for test reduction).  Ideally, the solution
should be implementable simply by modifying the predicate that is
evaluated.

A costly but effective solution is to make $N$ large in comparison to
the number of expected predicate evaluations performed during
delta-debugging.  If the error rate is low enough, the problem
disappears.  However, given the large number of evaluations performed,
this will tend to make delta-debugging extremely slow.  We propose
using a dynamic sampling approach, where $N$ is small, but if the
predicate evaluates to true, $M$ repeated true evaluations
(``replications'', where the first true evaluation is counted as 1 replication) are
required before the predicate returns ``true'' to the delta-debugging
algorithm.  A set of $M$ repeated false positives with
$\frac{N}{M}$ samples each is much less likely than a false positive
with $N$ samples; so long as we accept the resulting bias in favor of
false negatives, we can therefore produce a reduced test with a
desired $P(fail)$ much more cheaply:  the use of replications not only
means we only pay the full sampling price on rare occasions, but a
desired accuracy for $P$ can be obtained with a much smaller value $M
\times N$ than a non-dynamically-sampled $N$.  For example, if we the predicate
is $P(fail) \geq 0.5$, and the true probability for a candidate test
is 0.25, using $N=8$ will give a false positive rate of over 10\%.
Using 4 replications on just 2 samples ($M=4;N=2$) yields a false
positive rate of about 3.7\%, yet requires almost 60\% fewer
executions of the test.  The probability calculations for such comparisons
  are relatively simple, but in real testing are
  usually not very useful, since the true
  probability distributions of test behaviors vary widely and
  dynamically during the
  delta-debugging process, making exact calculation unwieldy, even if
  the changing probabilities were known; exact results would be based on fictional
  probabilities, so gathering experimental data during a trial
  delta-debugging run and tuning $N$ and $M$ to yield desired
  results is  more effective.

>>>>

Delta-debugging \cite{DD}\footnote{We use delta-debugging to stand for
  all test reduction algorithms, even those \cite{CReduce,tstl} that do not use the exact
  binary-search that distinguishes delta-debugging \emph{per se}; the
  differences are immaterial for our purposes.}  is a widely used method for reducing the
size of failing tests, making them easier to understand and debug.
The core idea of delta-debugging is to take a test that satisfies some
property (usually the property is ``it fails in a particular way'')
and produce a smaller test that satisfies the same property.
Delta-debugging, as proposed by Zeller and Hildebrandt, relies on a
modified binary search to produce this smaller test, though the
following discussion relies only on the general structure of
delta-debugging, and also applies to the greedy algorithms used to
reduce failing tests in some work \cite{CReduce,onetest}.  At a high
level delta-debugging and other test reduction/shrinking algorithms
can be seen as following a general algorithmic structure.  Given a
test case, $T_{\mathit{INIT}} : \mathit{test}$ and a predicate
$\mathit{PRED}: \mathit{test} \rightarrow \mathit{bool}$, we
\emph{reduce $T_{\mathit{INIT}}$ with respect to $\mathit{PRED}$} as follows:

@@

Delta-debugging in the context of detecting nondeterminism has two
purposes.  One is simply the usual goal of reducing the size of a
test.  Identifying the cause of nondeterminism may be very easy in a
test consisting of ten library function calls (it is one of these ten
calls), but very difficult in a test consisting of a hundred library
function calls.  This is no different than the common use of
delta-debugging.  However, in horizontal nondeterminism detection,
delta-debugging also tends to \emph{change the probability of
  nondeterministic behavior}; this can be both harmful and
beneficial.  The behavior is part of a more general (and, to our
knowledge, not previously investigated) issue:  using delta-debugging
to minimize a test with respect to a predicate $\mathit{PRED}$ that only holds with a
certain probability (the predicate itself is nondeterministic).  Note
that while most descriptions of delta-debugging discuss reducing a
test with respect to \emph{test failure} (hence the term
``delta-\emph{debugging}'') we adopt the wider view of delta-debugging
as reducing the length of a test while maintaining that an arbitrary
predicate remains true of the shorter versions of the test.  The
most common $\mathit{PRED}$ is still ``the test fails'' but other predicates,
especially those related to code-coverage, can also be useful
\cite{icst2014,stvrcausereduce,NonAdeq}.  Our contribution here is to consider
what happens when $\mathit{PRED}$ not deterministic.
``The test behaves flakily'' is a simple, important, example of such a $\mathit{PRED}$.

In \emph{monotonic} probabilistic delta-debugging, removing a part of a test cannot
increase the probability of the predicate holding.  This is fairly
common.  In these cases, when we use a reduction algorithm to reduce $t$
to $r$, $P(\mathit{PRED}(r)) \leq P(\mathit{PRED}(t))$.  In \emph{non-monotonic} cases,
however, there is no such upper bound.  Removing a step in $t$ may
increase the probability that $t$ behaves nondeterministically.

The $\mathit{PRED}$ of interest for horizontal determinism is basically of
the form:  {\bf the test will exhibit at least two different behaviors in $S$
runs, with probability $P$}.  This predicate may be monotonic or
non-monotonic, depending on the causes of the nondeterminism detected.

The only absolutely necessary changes required to use standard
delta-debugging implementations in nondeterminism detection
are simply removal of some ``sanity checks'' in the code:  many
implementations (including the Python code provided by Zeller) assert
that $\mathit{PRED}$ holds on the original test.  In probabilistic
settings, this may not be true, and we may even be trying to find a
subset where a predicate that is very far from holding on the original test
holds (in a non-monotonic case where we aim to increase the probability
of nondeterminism).

@@

However, simply using delta-debugging off the shelf with a $\mathit{PRED}$
such as $P(fail) > 0.3 \wedge P(fail) < 1.0$, to force a test to be flaky, and force it to fail
sufficiently often to be used in debugging, will often produce
surprising and unfortunate results.  In many cases,
delta-debugging will indeed reduce the large test to a small subset.
And, in a technical sense, delta-debugging will work:  it will never
convert a nondeterministic test to a completely deterministic test,
because the reduced test $r$ that delta-debugging returns is always one
such that $\mathit{PRED}(r)$ has evaluated to true at least once.
However, if you run the resulting test, it will, in many cases, have a
$P(fail)$ that is much, much smaller than
0.3, perhaps as low as 0.01.  Why?

The problem is analogous to the problem of publication bias in
scientific fields \cite{ahmed2012assessment}.  We can think of each
application of $\mathit{PRED}$ to a candidate variant
$T_{\mathit{NEW}}$ as a scientific experiment.  A predicate like
$P(fail) > 0.3 \wedge P(fail) < 1.0$ cannot be evaluated by
determining the true probability of failure, we have no more direct access to
this value than a medical researcher has access to the true effect
size and direction of a proposed treatment; rather, the test must be
run some concrete number of times, and the number of failures counted.
However, even if the number of samples $N$ is large, there is some
probability (based on the sample size) of a result that diverges
significantly from the actual probability of failure.  If the
predicate were run only once, and the number of samples reasonably
large, this would not matter.  However, by their nature,
delta-debugging and other test reduction algorithms explore a search
space that often contains hundreds or even thousands or millions of
candidate tests.  The predicate is evaluated on each of these, and so
even with large $N$, it is extremely likely that \emph{some} evaluation will
produce a very poor estimate of $P(fail)$.  If such an evaluation
causes the predicate to appear to hold for a test, delta-debugging is
``stuck'' with the error, because (to our knowledge) no test reduction
algorithms allow backtracking (in normal, non-probabilistic settings,
backtracking would not usually be useful).  After such a ``mistake,'' finding further
reductions will become harder, but may still be possible due to the
same source of errors:  \emph{the number of experiments run is far larger
  than the probability of an incorrect result for a single
  experiment}.  We define a \emph{false positive} in probabilistic
delta debugging as a case where $\mathit{PRED}$ evaluates to true, but
the actual probabilities required by $\mathit{PRED}$ do not hold, and
a false negative to be the reverse, where ``bad luck'' makes it appear
that the test does not satisfy $\mathit{PRED}$, but it does in fact do
so.  A false negative's consequences are limited, since there will
likely be another similar variation for which  $\mathit{PRED}$ will
evaluated to true, but a false positive changes $T_{\mathit{CURR}}$.

The analogy to publication bias is simple.  Assume that a paper in,
e.g., a medical journal will only be published if it shows that a
treatment is effective, with the statistical requirement that $p <
0.05$ (that is, roughly, by some statistical test, the probability
that the result is due to chance rather than reflecting reality, is
less than 5\%).  Even if no researchers are dishonest, 5 of every 100 papers published
in the journal will be report an effect with the wrong effect size or
even effect direction!  Now consider a ``hot'' topic, a kind of
treatment many research teams may investigate.  Major journals tend
not to publish papers that say ``this new treatment, which is not
established, may not work,'' (especially if the result is not ``the
treatment certainly does not work'' but only ``the experiment did not
establish $p < 0.05$''), but they  are very likely to publish a paper
that says ``this exciting new treatment works!''  In addition to
sometimes producing excitement about treatments that do not work, this
bias tends to magnify the expected effect size of treatments, since
the chance of getting a ``passing'' $p$-value is higher when an
experiment finds a larger effect size, due to how most statistical
tests work.
The \emph{bias} in
favor of positive results is shared by delta-debugging;
delta-debugging's behavior is much more influenced by cases where
$\mathit{PRED}$ holds than cases where it does not.  Looking at the
literature, the unpublished papers are invisible, magnifying the
effect on scientific opinion of the published experiments.

@@

In scientific literature, the most frequently proposed solution is the
use of replications:  repeated runs of ``successful'' experiments to
minimize the probability that a direction or effect size is a fluke due to publication
bias.  One way to produce this effect would be to allow
delta-debugging to backtrack if the probabilities observed in predicate
evaluations suddenly exhibit a strong discontinuity, a kind of ``paper
retraction'' based on near-replications.  However, this requires
modifying the delta-debugging implementations, which is difficult and
sometimes not really feasible (e.g., few users of afl-fuzz \cite{aflfuzz} will wish
to change its C code for test reduction).  Ideally, the solution
should be implementable simply by modifying the predicate that is
evaluated.

A costly but effective solution is to make $N$ large in comparison to
the number of expected predicate evaluations performed during
delta-debugging.  However, given the large number of evaluations performed,
this will tend to make delta-debugging extremely slow.  We propose
using a dynamic sampling approach, where $N$ is small, but if the
predicate evaluates to true, $M$ repeated true evaluations
(``replications'', where the first true evaluation is counted as 1 replication) are
required before the predicate returns ``true'' to the delta-debugging
algorithm.  To evaluate the predicate $\mathit{PRED}$ given $N$,
$M$, and desired probability threshold $p$, the algorithm is simple:

\begin{enumerate}
  \item  $i = 0$
  \item Evaluate $\mathit{PRED}$ $N$ times.  If the number of cases
  where $\mathit{PRED}$ was true is less than $p$, return false.
  \item  $i = i + 1$
  \item If $i < M$ go to step 1.
  \item  Return true.
\end{enumerate}

A set of $M$ repeated false positives with
$\frac{N}{M}$ samples each is much less likely than a false positive
with $N$ samples; so long as we accept the resulting bias in favor of
false negatives, we can therefore produce a reduced test with a
desired $P(fail)$ much more cheaply:  the use of replications not only
means we only pay the full sampling price on rare occasions, but a
desired accuracy for $P$ can be obtained with a much smaller value $M
\times N$ than a non-dynamically-sampled $N$.  For example, if $\mathit{PRED}$
is $P(fail) \geq 0.5$, and the true probability for a candidate test
is 0.25, using $N=8$ will give a false positive rate of over 10\%.
Using 4 replications on just 2 samples ($M=4;N=2$) yields a false
positive rate of about 3.7\%, yet requires almost 60\% fewer
executions of the test.  The probability calculations for such comparisons
  are relatively simple, but in real testing are
  usually not very useful, since the true
  probability distributions of test behaviors vary widely and
  dynamically during the
  delta-debugging process, making exact calculation unwieldy, even if
  the changing probabilities were known; exact results would be based on fictional
  probabilities, so gathering experimental data during a trial
  delta-debugging run and tuning $N$ and $M$ to yield desired
  results is  more effective.
